---
title: "Portfoliomanagement and Financial Analysis - Assignment 1"
subtitle: "Submit until Monday 2020-09-21, 13:00"
author: "Stefan, Macanovic"
output: html_notebook
---


```{r load_packs}
pacman::p_load(tidyverse,tidyquant,FFdownload)
```

For all exercises please refer to my book [Tidy Portfoliomanagement in R](https://bookdown.org/sstoeckl/Tidy_Portfoliomanagement_in_R). Unfortunately some functions (e.g. `get_exchange()`) are not working on rstudio.cloud, so please try to solve the exercises without. Also remember, that ecverything you do below must be **explained** to me (why? what does this mean? what is the interpretation?). If you have any questions or problems, please refer to the assignment forum on moodle.

## Exercise 1: xts

Create a daily time series for `r lubridate::year(Sys.time())`:

#1. Find the subset of first and last days per month/quarter (uniquely)

#First, we require the necessary packages.As followed:
require(zoo)
require(xts)
options(xts_check_TZ=FALSE)

#Definiton of TimeSequence
dates1 <-seq(as.Date("2020-01-01"), length=12, by="month")
#alternative 
dates2 <-timeSequence (from = "2020-01-01", to = "2020-12-31", by = "month")

#getting the information
timeFirstDayInMonth (dates1) #(dates1-7)-> first result=1.12.2019 but ends with 1.11.2020

timeFirstDayInQuarter(dates1)

timeLastDayInMonth(dates1)

timeLastDayInQuarter (dates1)


#2. Take December `r lubridate::year(Sys.time())-1` and remove all weekends and holidays in Zurich (Tokyo)

#first step - all days in December:
dates2 <- timeSequence (from = "2019-12-01", to = "2019-12-31", by = "days")
dates2

#second step - remove weekends - not necessary for the task
dates3 <- dates2[isWeekday(dates2)]
dates3

#third step - remove holidays
dates4 <- dates2[isBizday(dates2, holidayZURICH(2020))]
dates4 #we can see that on 25.12.2020 are holidays

#3. create a series of five dates & times in New York. Show them for New York, London and Belgrade

#GMT is starting zone
Dates <- c("2016-01-01", "2017-01-01", "2018-01-01", "2019-01-01", "2020-01-01")
Times <- c("15:22:02", "23:12:55", "10:34:02", "08:30:00", "11:18:23")
DatesTimes <- paste(Dates, Times)
as.Date(DatesTimes) 
NY<- timeDate(DatesTimes, zone= "GMT", FinCenter="NewYork") #turning GMT Date/Time to NYtime
NY
London <- timeDate(NY,zone="NewYork", FinCenter="London")
London
Belgrade <- timeDate (NY, zone="NewYork", FinCenter="Belgrade")
Belgrade


## Exercise 2: Download Financial Data and Merge

Download information on all stocks currently part of the S&P500. Select 10 that have data back to 2000. Additionally download the S&P 500 index and the four Fama-French factors including the risk-free rate (using my package [`ffdownload`](https://github.com/sstoeckl/ffdownload)). Merge all data together and calculate returns (simple or continuous returns?).
Split your data in two files: One holding daily returns and one with monthly returns. Additionally show how to we could get from long to wide format to transfer between `tibble` (`tsibble`) and xts.

```{r}
?FFdownload()
tempf <- tempfile(fileext = ".RData"); tempd <- tempdir(); temptxt <- tempfile(fileext = ".txt")
inputlist <- c("F-F_Research_Data_Factors","F-F_Momentum_Factor","F-F_ST_Reversal_Factor","F-F_LT_Reversal_Factor")
# Now process only these files if they can be matched (download only)
FFdownload(exclude_daily=TRUE,tempdir=tempd,download=TRUE,download_only=FALSE,inputlist=inputlist)
tempf2 <- tempfile(fileext = ".RData"); tempd2 <- tempdir() #asking Nicolas
FFdownload(output_file = tempf2,tempdir = tempd2,exclude_daily = TRUE, download = TRUE, download_only=FALSE, listsave=temptxt)
load(tempf2)
FFdownload$x_25_Portfolios_5x5$monthly$average_value_weighted_returns
```

```{r}
periodReturn
to.period()
```


#information on S&P 500

sp500 <-tq_index("SP500")

#daily

#Selection of 10 that have data back to 2000

startDate <- "2000-09-01"

stockSelectionDaily <- c('T', 'WMT', 'MSFT', 'BAC', 'PG', 'MCD', 'JPM', 'LIN', 'VZ', 'INTC') %>%

tq_get(get = "stock.prices", from = startDate, to = TODAY()) %>%

group_by(symbol) %>%
 
tq_transmute(select = adjusted, mutate_fun = periodReturn, period = "daily", col_rename =
'return')

sp500Daily <- tq_get("^GSPC", get = "stock.prices", from = startDate, to = TODAY()) %>%

tq_transmute(select = adjusted, mutate_fun = periodReturn, period = "daily", col_rename = 'SP500')

stockSelectionMonthly <- c('T', 'WMT', 'MSFT', 'BAC', 'PG', 'MCD', 'JPM', 'LIN', 'VZ', 'INTC') %>%

tq_get(get = "stock.prices", from = startDate, to = TODAY()) %>%
 group_by(symbol) %>%

tq_transmute(select = adjusted, mutate_fun = periodReturn, period = "monthly", col_rename =
'return')

sp500Monthly <- tq_get("^GSPC", get = "stock.prices", from = startDate, to = TODAY()) %>%

tq_transmute(select = adjusted, mutate_fun = periodReturn, period = "monthly", col_rename='SP500')

#preparing the data and merging it

require(timetk)

stockSelectionDaily_cols <- stockSelectionDaily %>% spread(symbol, return)

stockSelectionMonthly_cols <- stockSelectionMonthly %>% spread(symbol, return)

#stockSelectionDaily_xts <- tk_xts(stockSelectionDaily_cols, date_var = date)

dataDaily <- stockSelectionDaily_cols %>% merge(sp500Daily, by = "date")

dataMonthly <- stockSelectionMonthly_cols %>% merge(sp500Monthly, by = "date")

#fetch Fama-French data

tmpf <- tempfile(fileext = ".RData"); tmpd <- tempdir(); tmptxt <- tempfile(fileext = ".txt")

inputlist <- c("F-F_Research_Data_Factors","F-F_Momentum_Factor","F-F_ST_Reversal_Factor","FF_LT_Reversal_Factor")

FFdownload(output_file = tmpf, tempdir = tmpd, exclude_daily = FALSE, download = TRUE,
download_only = FALSE, listsave=tmptxt, inputlist=inputlist)

load(tmpf)

list.files(tmpd)

#prepare Fama-French data and merge it
ffFactorsDaily <- tk_tbl(FFdownload$`x_F-F_Research_Data_Factors`$daily$Temp2, timetk_idx =
TRUE, rename_index = "date")

ffFactorsDaily <- ffFactorsDaily[ffFactorsDaily$date >= startDate,,,,]
tail(ffFactorsDaily) #Fama-French data is only available till 30.06.2020 as of today
dataDaily <- dataDaily %>% merge(ffFactorsDaily, by = "date")
ffFactorsMonthly <- tk_tbl(FFdownload$`x_F-F_Research_Data_Factors`$monthly$Temp2,
timetk_idx = TRUE, rename_index = "date")%>%

mutate(date=as.Date(date, frac=1))
ffFactorsMonthly <- ffFactorsMonthly[ffFactorsMonthly$date >= startDate,,,,]
tail(ffFactorsMonthly) #Fama-French data is only available till 30.06.2020 as of today
dataMonthly <- dataMonthly %>% merge(ffFactorsMonthly, by = "date")

#export work into csv

write.table(dataDaily , file = "./dataDaily.csv")

write.table(dataMonthly, file = "./dataMonthly.csv")

#transform tibble into time series

install.packages("tsbox")

library(tsbox)

require(xts)

dailyTimeSeries <- ts_xts(dataDaily)

## Exercise 3: Plotting Data

Carefully work through `vignette("TQ04-charting-with-tidyquant")` and create plots using [ggplot](https://ggplot2.tidyverse.org/) from the tidyverse. Select four assets to highlight all functions that are important for you!

#install tidyverse and tidyquant package

install.packages("tidyverse")
install.packages("tidyquant")
install.packages("magrittr")

#get asset prices

getSymbols(Symbols = "AMZN", from="2019-09-01", to="2020-09-01", periodicity="monthly")
getSymbols(Symbols = "AAPL", from="2019-09-01", to="2020-09-01", periodicity="monthly")
getSymbols(Symbols = "MSFT", from="2019-09-01", to="2020-09-01", periodicity="monthly")
getSymbols(Symbols = "GOOG", from="2019-09-01", to="2020-09-01", periodicity="monthly")

stocks <- as.xts(data.frame(AAPL = AAPL[, "AAPL.Close"], AMZN = AMZN[, "AMZN.Close"],
 MSFT = MSFT[, "MSFT.Close"], GOOG = GOOG[, "GOOG.Close"]))
 
head(stocks)

#plotting the data

plot(as.zoo(stocks), screens = 1, lty = 1:4, xlab = "Date", ylab = "Price")
legend("right", c("AAPL", "AMZN", "MSFT", "GOOG"), lty = 1:4, cex = 0.5)

#convert xts into matrix
vecst<-data.matrix(as.data.frame(stocks)) #convert xts into matrix

#boxplot of matrix
boxplot(vecst)

#historgram with normal curve
h<-hist(vecst, breaks=10, col="red", xlab="ClosingPrice",
 main="Histogram with Normal Curve")
 xfit<-seq(min(vecst),max(vecst),length=40)
yfit<-dnorm(xfit,mean=mean(vecst),sd=sd(vecst))
yfit <- yfit*diff(h$mids[1:2])*length(vecst)
lines(xfit, yfit, col="blue", lwd=2)

#normal qq-plot
qqnorm(stocks)

```{r}
ggplot(aes()) + ... + facet_wrap()
```

Additionally we need statistical information, so we want to plot a histogram (including normal density), a box-plot and a qq-plot. Try to accumulate data in a meaningful way to not have 10 different plots!

## Exercise 4: Statistics

#Here you might use all `tidyquant`-vignettes! My shiny-script here might also help you to get an idea: [www.riskindex.eu:3838/mci](www.riskindex.eu:3838/mci).

### Exercise 4.1 Within securities

#For the four securities calculate all interesting statistics (measures of location, dispersion, skewness and kurtosis). Check returns for univariate (and multivariate) normality.

```{r}
table.Stats()
table.DownsideRisk
table.HigherMoments
```

#Require packages
require(tidyverse)
require(tidyquant)
require(timetk)
options(xts_check_TZ=FALSE)

#Data download
Stocks.raw <- c("MSFT","AAPL","AMZN","GOOG") %>%
    tq_get(get  = "stock.prices", from = "2000-01-01", to = Sys.Date()) %>%
    group_by(symbol)
Stocks_ret_daily <- Stocks.raw %>%
    tq_transmute(select = adjusted, mutate_fun = periodReturn,
                 period = "daily", col_rename = "return", type="log")
Stocks_ret_monthly <- Stocks.raw %>%
    tq_transmute(select = adjusted, mutate_fun = periodReturn,
                 period = "monthly", col_rename = "return", type="log")

#geometric and arithmetic mean

#monthly
Stocks_ret_monthly %>% summarize(geometric_mean=mean.geometric(return), arithmetic_mean=mean(return)) %>%
          mutate_at(-1, round, 5) %>% t()
          
#daily
Stocks_ret_daily %>% summarize(geometric_mean=mean.geometric(return), arithmetic_mean=mean(return)) %>%
          mutate_at(-1, round, 5) %>% t()

#Skewness and kurtosis

#monthly
Stocks_ret_monthly %>% summarize(Skewness=skewness(return), "Ex. Kurtosis"=kurtosis(return)) %>%
          mutate_at(-1, round, 3) %>% t()

#daily
Stocks_ret_daily %>% summarize(Skewness=skewness(return), "Ex. Kurtosis"=kurtosis(return)) %>%
          mutate_at(-1, round, 3) %>% t()
          
#measures of spread

#monthly
Stocks_ret_monthly %>% summarize(Volatility=sd(return), Range=max(return)-min(return), IQR=IQR(return)) %>%
          mutate_at(-1, round, 3) %>% t()
          
#daily
Stocks_ret_daily %>% summarize(Volatility=sd(return), Range=max(return)-min(return), IQR=IQR(return)) %>%
          mutate_at(-1, round, 3) %>% t()

#Standard deviation - measures of spread financial risk

#monthly
tocks_ret_monthly %>% summarize(Semideviation=SemiDeviation(return), 
  "VaR(0.05)"=VaR(return,p=0.05), "CVaR(0.05)"=CVaR(return,p=0.05)) %>%
  mutate_at(-1, round, 3) %>% t()

#daily
Stocks_ret_daily %>% summarize(Semideviation=SemiDeviation(return), 
  "VaR(0.05)"=VaR(return,p=0.05), "CVaR(0.05)"=CVaR(return,p=0.05)) %>%
  mutate_at(-1, round, 3) %>% t()


#correlation

#monthly
chart.Correlation(Stocks_xts_monthly, histogram=TRUE, pch="+")

#daily
chart.Correlation(Stocks_xts_daily, histogram=TRUE, pch="+")


### Exercise 4.2 Between securities and with index

#Plot all securities vis-a-vis the index. Include a regression line and regression information. Can you plot rolling correlations between assets (and/or the index)?

#grouping the 4 securities, getting monthly returns
Ra<-c("AAPL","AMZN","MSFT","GOOG")%>%
tq_get(get="stock.prices",from="2016-12-31",to="2019-12-31")%>%
group_by(symbol)%>%
tq_transmute(select     = adjusted, mutate_fun = periodReturn,period     = "monthly",col_rename = "Ra")

#monthly Returns for the Dow Jones
Rb <- c("DJI") %>%
    tq_get(get  = "stock.prices",from="2016-12-31",to="2019-12-31")  %>%
    tq_transmute(select     = adjusted,
                 mutate_fun = periodReturn,
                 period     = "monthly",
                 col_rename = "Rb")
                 
RaRb <- left_join(Ra, Rb, by = c("date" = "date"))

RaRb %>%
    tq_performance(Ra = Ra, Rb = Rb, performance_fun = table.CAPM)

#AAPL vs Dow Jones    
stock_prices <- c("DJI", "AAPL") %>%
    tq_get(get  = "stock.prices",
         from="2016-12-31",to="2019-12-31") %>%
    group_by(symbol)

stock_pairs <- stock_prices %>%
    tq_transmute(select     = adjusted,
                 mutate_fun = periodReturn,
                 period     = "daily",
                # type       = "log",
                 col_rename = "returns") %>%
    spread(key = symbol, value = returns)

stock_pairs %>%
    ggplot(aes(x = DJI, y = AAPL)) +
    geom_point(color = palette_light()[[1]], alpha = 0.5) +
    geom_smooth(method = "lm") +
    labs(title = "Returns Relationship of Apple and Dow Jones 2016/12/31 to 2019/12/31 - Method: lm") +
    theme_tq()

#AMZN vs Dow Jones
stock_prices <- c("DJI", "AMZN") %>%
    tq_get(get  = "stock.prices",
         from="2016-12-31",to="2019-12-31") %>%
    group_by(symbol)

stock_pairs <- stock_prices %>%
    tq_transmute(select     = adjusted,
                 mutate_fun = periodReturn,
                 period     = "daily",
                # type       = "log",
                 col_rename = "returns") %>%
    spread(key = symbol, value = returns)

stock_pairs %>%
    ggplot(aes(x = DJI, y = AMZN)) +
    geom_point(color = palette_light()[[1]], alpha = 0.5) +
    geom_smooth(method = "lm") +
    labs(title = "Returns Relationship of Amazon and Dow Jones 2016/12/31 to 2019/12/31 - Method: lm") +
    theme_tq()
    
#MSFT vs Dow Jones
stock_prices <- c("DJI", "MSFT") %>%
    tq_get(get  = "stock.prices",
         from="2016-12-31",to="2019-12-31") %>%
    group_by(symbol)

stock_pairs <- stock_prices %>%
    tq_transmute(select     = adjusted,
                 mutate_fun = periodReturn,
                 period     = "daily",
                # type       = "log",
                 col_rename = "returns") %>%
    spread(key = symbol, value = returns)

stock_pairs %>%
    ggplot(aes(x = DJI, y = MSFT)) +
    geom_point(color = palette_light()[[1]], alpha = 0.5) +
    geom_smooth(method = "lm") +
    labs(title = "Returns Relationship of Microsoft and Dow Jones 2016/12/31 to 2019/12/31 - Method: lm") +
    theme_tq()
    
#Google vs Dow Jones

stock_prices <- c("DJI", "GOOG") %>%
    tq_get(get  = "stock.prices",
         from="2016-12-31",to="2019-12-31") %>%
    group_by(symbol)

stock_pairs <- stock_prices %>%
    tq_transmute(select     = adjusted,
                 mutate_fun = periodReturn,
                 period     = "daily",
                # type       = "log",
                 col_rename = "returns") %>%
    spread(key = symbol, value = returns)

stock_pairs %>%
    ggplot(aes(x = DJI, y = GOOG)) +
    geom_point(color = palette_light()[[1]], alpha = 0.5) +
    geom_smooth(method = "lm") +
    labs(title = "Returns Relationship of Google and Dow Jones 2016/12/31 to 2019/12/31 - Method: lm") +
    theme_tq()


#Can you plot rolling correlations between assets (and/or the index)?
Ra <- c("AAPL", "GOOG", "MSFT","AMZN") %>%
    tq_get(get  = "stock.prices",
          from = "2013-12-31",
           to   = "2018-12-31") %>% 
    group_by(symbol) %>%
    tq_transmute(select     = adjusted,
                 mutate_fun = periodReturn,
                 period     = "monthly",
                  type       = "log",
                 col_rename = "Ra")

Rb <- c("DJI") %>%
    tq_get(get  = "stock.prices",
           from = "2013-12-31",
           to   = "2018-12-31") %>%
    tq_transmute(select     = adjusted,
                 mutate_fun = periodReturn,
                 period     = "monthly",
                  type       = "log",
                 col_rename = "Rb")

RaRb <- left_join(Ra, Rb, by = c("date" = "date"))

rolling_corr <- RaRb %>%
    tq_transmute_xy(x          = Ra,
                    y          = Rb,
                    mutate_fun = runCor,
                    n          = 6,
                    col_rename = "rolling.corr.6")

rolling_corr %>%
    ggplot(aes(x = date, y = rolling.corr.6, color = symbol)) +
    geom_hline(yintercept = 0, color = palette_light()[[1]]) +
    geom_line(size = 1) +
    labs(title = "Six Month Rolling Correlations to Dow Jones (2013/12/31 to 2018/12/31)",
         x = "", y = "Correlation", color = "") +
    facet_wrap(~ symbol, ncol = 2) +theme_tq() +
    scale_color_tq()





#####entwurf


```{r}
table.CAPM
```

## Exercise 5: Performance Analysis

Create multiple portfolios using `tq_portfolio()` (e.g. an equally weighted one, a market-value weighted one, and one where you create a trading strategy according to your liking). DO a performance analysis a la `vignette("TQ05-performance-analysis-with-tidyquant")`.


To make a portfolio I have to choose assets. Therefore I took the monthly stock prices from Apple, Google, Netflix and Amazon from 31.12.2013 to 31.12.2019.
```{r}
stock_returns_monthly <- c("AAPL", "GOOG", "NFLX", "AMZN") %>%
  tq_get(get  = "stock.prices",
         from = "2013-12-31",
         to   = "2019-12-31") %>%
  group_by(symbol) %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "monthly", 
               col_rename = "Ra")
```

Now I create the repetition for 3 portfolios
```{r}
stock_returns_monthly_multi <- stock_returns_monthly %>%
  tq_repeat_df(n = 3)
```

#For the market-value weighted portfolio I get the market values as of 13 September 2019 (Source:ycharts.com) for portfolio 2
```{r}
MV_AAPL<-988.57
MV_GOOG<-859.51
MV_NFLX<-128.79
MV_AMZN<-909.84

MV_sum<-MV_NFLX+MV_AAPL+MV_AMZN+MV_GOOG
```

#Now I create the vector for the equally weighted one (25% for each stock). One portfolio is equally weighted with 25% and the 3rd portfolio is weighted after my preferences
```{r}
weights <- c(
  0.25, 0.25, 0.25, 0.25, 
  MV_AAPL/MV_sum, MV_GOOG/MV_sum, MV_NFLX/MV_sum, MV_AMZN/MV_sum, 
  0.2, 0.4, 0.2, 0.2)

```

#To measure the performance of the 3 portfolios I choose the SP500 as baseline returns
```{r}
baseline_returns_monthly <- "^GSPC" %>%
  tq_get(get  = "stock.prices",
         from = "2013-12-31",
         to   = "2019-12-31") %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "monthly", 
               col_rename = "Rb")
```


#The next step is to create the portfolios according to the weight table
```{r}
stocks <- c("AAPL", "GOOG", "NFLX", "AMZN")
weights_table <-  tibble(stocks) %>%
  tq_repeat_df(n = 3) %>%
  bind_cols(tibble(weights)) %>%
  group_by(portfolio)
```

#See the weights:
```{r}
weights_table
```

#The next step is to create the returns for the portfolios. 
```{r}
portfolio_returns_monthly_multi <- stock_returns_monthly_multi %>%
  tq_portfolio(assets_col  = symbol, 
               returns_col = Ra, 
               weights     = weights_table, 
               col_rename  = "Ra")
portfolio_returns_monthly_multi
```

#Combine the portfolios and baseline
```{r}
RaRb_multiple_portfolio <- left_join(portfolio_returns_monthly_multi, 
                                     baseline_returns_monthly,
                                     by = "date")
```

#Now I can make a performance analysis. Therefore I create Table CAPM
```{r}
RaRb_multiple_portfolio %>%
  tq_performance(Ra = Ra, Rb = Rb, performance_fun = table.CAPM)
```

#Table Sharpe Ratio
```{r}
RaRb_multiple_portfolio %>%
  tq_performance(Ra = Ra, Rb = NULL, performance_fun = SharpeRatio)
```

#Create Table Stats
```{r}
RaRb_multiple_portfolio %>%
  tq_performance(Ra = Ra, Rb = NULL, performance_fun = table.Stats)
```